### 2021-12-26 paper.tex
task 2

%===========================================================
\section{Task 3}
Focus tracker accumulates evidence of a changing state over the course of the dialogue. Let's define for methods and goal components $c \in V_c$ (all possible slot values) the quantity denoting that a slot(method) is not present in turn $t$:

@@

    For requestable slots.
    Assuming that probabilities are well-defined for $t = k-1$:
    \begin{equation}
        0 \leq p_{s,k-1} \leq 1
    \end{equation}
    Since
    \begin{equation}
        0 \leq slu_{s,t} + (1 - slu_{s,t}) \leq 1
    \end{equation}

@@

\subsection{Exercise 4}

$slu = slu_{c, 1}(v) = slu_{c,2}(v)$ for $\forall c \forall v \in V_c \Rightarrow q_{c,1} = q_{c, 2} = q$

@@

So the probability of value 'none' being 'none' when it is repeated twice decreases.

\subsection{Exercise 5}

\begin{equation}
    p_{c,t}(v) = (1-q_{c,t})\cdot slu_{c,t}(v) + q_{c,t} \cdot p_{c,t-1}(v)
\end{equation}

Since $q_{c,0}=1$ means that all ptobabilities were 0 for any $v$ apart from 'none' value, so $p_{c,0}(v) = 0$ for all $v \in V_c$

Then $p_{c,1}(v) = (1 = 0.5) \cdot 0.5 + 0 = 0.25$

As the slot was not in the first turn and then appeared in the second turn with probability 0.5, then it is okay to assume that the the probability of it might be not 0.5 (as it would be in previous update rule) but 0.25.

In general, such rule puts less weights on the new probabilities from current turn and puts more weights on the previouos turn than before.

\subsection{Exercise 3}
it performes much better for the informable slots. This is an expected behaviour as informable slot may change over time, and the probability of the slot being informable will change over time. However for the requestable slot focus tracker doesn't show a better performance and usually the requestable slots do not change during the dialog, they are requested once and after that removed from the set of the requestable slots. The same applied for the methods, performance there is the similar to the baseline tracker as user only once "sets" by which method he wants to achieve the goal.
\section{Monte-Carlo policy}

@@

It is a tabular approach, which means that the belief state/summary space is discretised into a grid. In this algorithm the policy and the Q function are first defined randomly. Then we iterate over the dialogs with this polilcy. Throught this iteration we calculate the reward until the end of the dialog. For updating the state we choose the nearest state from the table and update the Q value and afterwards the policy with it, then in the next turn we will use an already updated policy.

>>>>

\pagebreak
\section{Task 2 Semantic Decoding}
Semantic decoding translates sentences into the set of sematic concepts, which consists of 2 parts: intent (or dialog act type) and sematic slots and values. For instance the sentence "I want to book a flight to Berlin" can be translated into the intent "book" and the slots "flight" and "to" with the value "Berlin"($book(flight, to(Berlin))$).

To do this we first need to translate a sentence into a sequence of tags, for ex. "I want to fly from Minsk to Kiev round trip tomorrow" can be translated into the sequence of tags: ["O", "O", "O", "O", "O", "B-fromloc.city\_name", "O", "B-toloc.city\_name", "B-round\_trip", "I-round\_trip", "B-departure\_date"].

In order to do this, we can keep two dictionaries (one for words and one for tags), which will map each word (tag) to a unique integer. We can then use these dictionaries to translate the sequence of words and a sequence of tags into a sequence of integers. Then we can train a model to predict from the sequence of integers a corresponding sequence of tags. This can be a simple RNN with two heads - one for predicting the tags and another for predicting intent.

There are different ways of how we can evaluate such model. But first, having a sequence of tags, we can convert it to a set of slots in the following way: input = [B-food-type I\_food-type O B\_price-range O], output = [(food-type, start=0, end=1), (price-range, start=3, end=3)]. Then having a set of true and predicted slots we can compare them on the slot level: if predicted slot was in true slots, then this is a true positive. If predicted slot was not in true slots, then this is a false positive. If true slot was not in predicted slots, then this is a false negative. Then from these values, we can calculate precision, recall and F1 score.

On the other hand, we can calculate accuracy directly on the predicted sequence of tags, without converting it to slots. In this case we keep the mapping of each tag to 3 counters for TP, FP and FN. After that there are 2 options on how to average the metrics. On a macro level, we first calculate precision, recall and f1-score for each of the tags, then we average these metrics across all tags. On a micro level we do the weighted average, so for ex. to calculate precision we divide the sum of all TP by the sum of all TP+FP of all tags.

Macro average can be much lower than a weighted average, because there can be a lot of rare classes with few instances which were not predicted or predicted wrongly while there are not so many classes with a lot of instances, which are therefore predicted correctly.
%===========================================================
\section{Task 3}
Focus tracker accumulates evidence of a changing state over the course of the dialogue. Let's define for methods and goal components $c \in V_c$ (all possible slot values) the quantity denoting that a slot(method) is not present in turn $t$:

@@

    For requestable slots.\\

    Assuming that probabilities are well-defined for $t = k-1$:
    \begin{equation}
        0 \leq p_{s,k-1} \leq 1
    \end{equation}

    Since

    \begin{equation}
        0 \leq slu_{s,t} + (1 - slu_{s,t}) \leq 1
    \end{equation}

@@

\subsection{Exercise 3}
Focus tracker performes much better for the informable slots. This is an expected behaviour as informable slot may change over time, and the probability of the slot being informable will change over time. However for the requestable slot focus tracker doesn't show a better performance and usually the requestable slots do not change during the dialog, they are requested once and after that removed from the set of the requestable slots. The same applied for the methods, performance there is the similar to the baseline tracker as user only once "sets" by which method he wants to achieve the goal.

\subsection{Exercise 4}
What happens to a probability if a user gives the exact same utterance in
the first two turns of the dialog?

$slu = slu_{c, 1}(v) = slu_{c,2}(v) \Rightarrow q_{c,1} = q_{c, 2} = q$ for $\forall c \forall v \in V_c $

@@

So the probability of value 'none' when the same utterance is repeated twice decreases.

\subsection{Exercise 5}
Let's defince another update rule:
\begin{equation}
    p_{c,t}(v) = (1-q_{c,t})\cdot slu_{c,t}(v) + q_{c,t} \cdot p_{c,t-1}(v)
\end{equation}
Assume we talked to the user for two turns and for an informable slot $c$ it holds that 
\begin{equation}
    \label{eqn:q_c_0_1}
    q_{c, 0} = 1
\end{equation}
and 
\begin{equation}
    \label{eqn:q_c_1}
    q_{c, 1} = slu_{c,1}(v) = 0.5
\end{equation}

\noindent What happens to the probability $p_{c,1}(v)$?\\

\noindent Since (\ref{eqn:q_c_0_1}) means that all probabilities were 0 for any $v$ apart from 'none' value, so $p_{c,0}(v) = 0$ for all $v \in V_c$ \\

\noindent Then $p_{c,1}(v) = (1 = 0.5) \cdot 0.5 + 0 = 0.25$ \\

\noindent As the slot was not in the first turn and then appeared in the second turn with probability 0.5, then it is okay to assume that the the probability of it might be not 0.5 (as it would be in previous update rule) but 0.25. \\

\noindent In general, such rule puts less weight on the new probabilities from current turn and more weight on the previous turn, which is also a totally reasonable update.

\section{Monte-Carlo policy}

@@

\noindent It is a tabular approach, which means that the belief state/summary space is discretised into a grid. In this algorithm the policy and the Q function are first defined randomly. Then we iterate over the dialogs with this polilcy. Throught this iteration we calculate the reward until the end of the dialog. For updating the state we choose the nearest state from the table and update the Q value and afterwards the policy with it, then in the next turn we will use an already updated policy.
